---
output:
  word_document: default
  html_document: default
---
## BAN502 Module4


## Assignment 1. Missing Data


## Mahepath-Corbitt, Niesha


```{r}
options(tidyverse.quiet = TRUE)
library(tidyverse)
library(VIM)
library(mice)
grades <- read_csv("class-grades.csv")
```


Task 1: How much data is missing and in what variables? 
There are 11 NAs in the data: 1 in Tutorial, 3 in Midterm, 3 in TakeHome, and 4 in Final. 
```{r}
str(grades)
summary(grades)
```

Task 2: Use the VIM package to visualize missingness. Does there appear to be systematic missingness? In other words, are there students that are mising multiple pieces of data?
Yes. There are students missing several data in midterms, takehome, and final, but most of the data is represented in the Final row with 4 variables.
```{r}
vim_plot = aggr(grades, numbers = TRUE, prop = c(TRUE, FALSE), cex.axis = .7)
```


Task 3: Use row-wise deletion of missing values to create a new data frame. How many rows remain in this data frame?
After using drop_na function to do a row-wise deletion on the Final variable, there are now 95 variables (rows) left in the new data frame, grades1.
```{r}
gradesrow = grades %>% drop_na(Final)
str(gradesrow)
```


Task 4: Use column-wise deletion of missing values to create a new data frame (from the original data frame not from the data frame created in Task 3). How many columns remain in this data frame?
Using the column-wise deletion to delete the Final row, has completely deleted the column and now there are only 89 variable left in the new data frame, gradescolumn.
```{r}
gradescolumn = grades %>% select(-Final)
vim_plot1 = aggr(gradescolumn, number = TRUE, prop = c (TRUE, FALSE), cex.axis = .7)
```


Task 5: Which approach (Task 3 or Task 4) seems preferable for this dataset? Brieﬂy discuss your answer
As more important data has been removed with the column-wise deletion, the row-wise deletion is the more preferable method.

Task 6 Use the code below to impute the missing values in the dataset using the mice package. 

```{r}
grades_imp = mice(grades, m=1, method = "pmm", seed = 12345) #in line above: m=1 -> runs one imputation, seed sets the random number seed to get repeatable results 
summary(grades_imp) 
densityplot(grades_imp) #red imputed, blue original, only shows density plots when more than 1 value the variable was imputed #note that the density plots are fairly uninteresting given the small amount of missing data 
grades_complete = complete(grades_imp) 
summary(grades_complete)
```


Task 7: Brieﬂy discuss potential issues that could be encountered when working with missing data. Describe situations where imputation may not be advisable.

Imputation may not be advisable when the data is skewed.